Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-09-28 grove-paper.tex
Rough draft of Section 2.3.7 Disconnection

% math ligatures
\mathlig{::=}{\coloncolonequals}
\mathlig{|-}{\vdash}
\mathlig{:->}{\mapsto}
\mathlig{-->a}{\overset{a}{\longrightarrow}}
\mathlig{||}{|}
\mathlig{|}{:}
\mathlig{:}{\!:\!}

@@

@@

NEEDS UI MOCKUP

(Maybe not in initial grove by example) The A and B diamond example

@@

\begin{figure*}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{1}_{4}$" [> "left", > "19"'],
   },
   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
    "$\texttt{2}_{4}$" [> "left", > "19"'],
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{1}_{4}$" [> "left", > "19"'],
    "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
     "$\texttt{2}_{4}$" [> "left", > "19"'],
    }
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
    "$\texttt{2}_{4}$" [> "left", > "19"'],
    "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
     "$\texttt{1}_{4}$" [> "left", > "19"'],
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-c}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"']
 }
};
\path (-1cm,-2cm) graph[graph style] {
 "$\texttt{+}_{6}$" [> "left", > "19"'] -> {
  "$\texttt{2}_{4}$" [> "left", > "19"'],
 }
};
\path (1cm,-2cm) graph[graph style] {
 "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
  "$\texttt{1}_{4}$" [> "left", > "19"'],
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-d}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Disconnection}
\label{fig:Disconnection}
\end{figure*}

@@

\begin{align*}
        \vertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \edges(G) \}               \\
    \liveVertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}           \\
 \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
         \orphans(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || = 0 \}       \\
    \multiparents(G) &= \{ v \in \liveVertexes(G) | || \parents(G,v) || > 1 \}       \\
        \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
\end{align*}

>>>>

% math ligatures
\mathlig{::=}{\coloncolonequals}
\mathlig{|-}{\vdash}
\mathlig{:->}{\mapsto}
\mathlig{-->a}{\overset{a}{\longrightarrow}}
%\mathlig{||}{|}
%\mathlig{|}{:}
%\mathlig{:}{\!:\!}
\newcommand{\abs}[1]{\left|{#1}\right|}

@@

TODO: NEEDS UI MOCKUP

@@

@@

\begin{figure*}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'],
   "$\texttt{*}_{6}$" [> "left", > "19"']
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{*}_{6}$" [> "left", > "19"'],
    {}
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{*}_{6}$" [> "left", > "19"'] -> {
    "$\texttt{+}_{2}$" [> "left", > "19"'],
    {}
   }
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-c}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"']
 }
};
\path (-1cm,-3cm) graph[graph style] {
 times/"$\texttt{*}_{6}$" [> "left", > "19"']
};
\path (1cm,-3cm) graph[graph style] {
 plus/"$\texttt{+}_{2}$" [> "left", > "19"']
};
\path[draw,->] (times) to [out=45,in=135] (plus);
\path[draw,->] (plus) to [out=-135,in=-45] (times);
\end{tikzpicture}
\caption{Initial code}
\label{fig:Disconnection-d}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Disconnection}
\label{fig:Disconnection}
\end{figure*}

Finally, when multi-user edits are merged,
parts of the graph can become disconnected
even though they are connected in each user's
copy of the graph before the merge.

For example, consider the situation in \autoref{fig:Disconnection}.
\autoref{fig:Disconnection-a} shows the graph after Alice and Bob have made some more edits,
and Alice and Bob's editors are synchronized with each other.
This graph represents the expression~\texttt{(\hole+\hole)+(\hole*\hole)}.
Alice moves the \texttt{*} under \texttt{+}$_2$.
This results in \autoref{fig:Disconnection-b},
which represents the expression~\texttt{((\hole*\hole)+\hole)+\hole}.
On the other hand, Bob puts \texttt{+}$_2$ and under the \texttt{*}.
This results in \autoref{fig:Disconnection-c},
which represents the expression~\texttt{\hole+((\hole+\hole)*\hole)}.

When these edits are merged, the result is the graph in \autoref{fig:Disconnection-d}.
Since Alice deleted the edge from \Vertex{TODO} to \Vertex{TODO}
and Bob deleted the edge from \Vertex{TODO} to \Vertex{TODO},
when these are merged \Vertex{TODO} and \Vertex{TODO}
are completely disconnected from the root.

This is similar to the type of merge conflict
that can arise in traditional version control systems
when two users delete different functions that
are duplicates of the other.

Fortunately, we can detect this situation and warn the user
since the disconnection appears after merging two graphs
that did not have disconnected vertexes.

@@

\begin{align*}
        \vertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \edges(G) \}               \\
    \liveVertexes(G) &= \{ v \in \V | \exists (c,v,u) \in \liveEdges(G) \}           \\
 \childVertexes(G,v) &= \{ v' \in \V | \exists (c,v',u) \in \vertexChildren(G,v) \}  \\
         \orphans(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} = 0 \}       \\
    \multiparents(G) &= \{ v \in \liveVertexes(G) | \abs{\parents(G,v)} > 1 \}       \\
        \vertex(G,u) &= v \in \vertexes(G) \iff (\exists k \in \K)(v = (k,u))
\end{align*}
